---
output:
  html_document: default
  pdf_document: default
---
***Energy Star Score Prediction for NY Buildings***

**Data Importing**
```{r, message=FALSE, results='hide'}
# Importing the required libraries
library(Hmisc)
library(caret)
library(corrplot)
library(Metrics)

memory.limit()
memory.limit(size=16000)
```

```{r}
#Importing the dataset
data <- read.csv(file = 'Usecase1_Dataset.csv')
#summary (data)
dim(data)
```
**Data Cleaning and Preprocessing**

```{r}
#There are 60 variables including a mix of discrete and continuous variables

#We can remove the unwanted columns like additional location details (Street number, street address) which is taken care by Postal Code variable

#Assumption 1 - Property id and other property details do not affect the energy score as these are only identification details for record purposes

#Assumption 2 - Duplicate entries like 'Primary Property Type' can be omitted since the data is covered well in 'Largest property use type'

data_clean <- data[,11:60]
data_clean$Street.Number <- NULL
data_clean$Street.Name <- NULL
data_clean$List.of.All.Property.Use.Types.at.Property <- NULL
data_clean$Primary.Property.Type...Self.Selected <- NULL
```

```{r}
#Postal code and borough seem to be correlated
#There are more missing values in Borough than in postal code,hence dropping Borough column
data_clean$Borough<- NULL
data_clean$NTA <- NULL
data_clean$DOF.Gross.Floor.Area <- NULL
data_clean$Release.Date <- NULL
```

```{r}
#Removing latitude & longitude columns as we have postal code to group according to area
data_clean <- data_clean[,1:(ncol(data_clean)-5)]

#Tackling rows with 'Not available' data for energy score (target variable)
#Assumption 3 - Removing data points with NaN values of target variable will not affect the accuracy of model as the removed data points are much less
sum(data_clean$ENERGY.STAR.Score == "Not Available")
data_clean <- data_clean[-which(data_clean$ENERGY.STAR.Score == "Not Available"),]
```

```{r}
#Dropping variables not normalized for weather and as we are taking (/ftA2) values instead of total value
data_clean$`Site.EUI..kBtu.ft².` <- NULL
data_clean$`Source.EUI..kBtu.ft².` <- NULL
data_clean$Total.GHG.Emissions..Metric.Tons.CO2e. <- NULL
data_clean$Weather.Normalized.Site.Electricity..kWh. <- NULL
data_clean$Water.Required. <- NULL
data_clean$Water.Use..All.Water.Sources...kgal. <- NULL
```

**Remove data columns with more than 40% of 'Not Available' entries**
```{r}
cnt_NA <- rep(0,31)

for (i in 1:31){
  if (length(grep("Not Available", data_clean[,i])) > 0){
    cnt_NA[i] <- length(which(data_clean[,i] == "Not Available")) 
  }
}
rm(i)

percent_NA <- (cnt_NA/nrow(data_clean))*100
df_NAcolumns <- cbind(colnames(data_clean), percent_NA)
rm(cnt_NA)

data_clean <- data_clean[,which(percent_NA <= 40)]
rm(percent_NA)
```


```{r}
#Arrange columns per factored, continuous and target variable/s
data_clean <- subset(data_clean,select = c(1,2,7,8,21,3:6,10:20,9))
```

```{r}
#Converting Postal code data to valid data, truncating first five digits (as is the standard format)
data_clean$Postal.Code <- substr(data_clean$Postal.Code, 1,5)
```

```{r, warning=FALSE}
#Convert the variables to factors and numericals
for (i in 1:5){
  data_clean[,i] <- as.factor(data_clean[,i])
}

for (i in 6:21){
  data_clean[,i] <- as.numeric(data_clean[,i])
}
rm(i)

summary(data_clean)
```

**Understanding class distribution of levels for discrete variables**
**As can be seen in the proportion table, 'Metered areas Energy' column has skewed ratio of the factors**
**and very less data points for 'another config' and 'not available', hence removing these rows (20 count) and the column**
**Similarly with 'Metered areas water', removing the rows with factors of less data points and removing the column**

```{r}
#Bar Plot for Discrete Variables
par(mfrow=c(2,2))
barplot(prop.table(table(data_clean$Metered.Areas..Energy.)), ylab = "Frequency", 
        main = "Frequency chart for 'Metered Areas Energy'", col = "blue")
barplot(prop.table(table(data_clean$Metered.Areas...Water.)), ylab = "Frequency", 
        main = "Frequency chart for 'Metered Areas Water'", col = "Cyan")
barplot(prop.table(table(data_clean$Largest.Property.Use.Type)), ylab = "Frequency", 
        main = "Frequency chart for 'Largest.Property.Use.Type'", col = "Red")
barplot(prop.table(table(data_clean$Postal.Code)), ylab = "Frequency", 
        main = "Frequency chart for 'Postal.Code'", col = "Black")
par(mfrow=c(1,1))
```

```{r}
#Cleaning Metered Areas Energy column
data_clean <- data_clean[which(data_clean$Metered.Areas..Energy. == "Whole Building"),]
data_clean$Metered.Areas..Energy. <- NULL
```

```{r}
#Cleaning 'Metered Area Water' column
data_clean <- data_clean[which(data_clean$Metered.Areas...Water. == "Not Available" | data_clean$Metered.Areas...Water. == "Whole Building"), ]
data_clean$Metered.Areas...Water. <- NULL
```

```{r}
#Cleaning 'Largest Property Use Type' column
table(data_clean$Largest.Property.Use.Type )

data_clean <- data_clean[which(data_clean$Largest.Property.Use.Type == "Multifamily Housing" |
                                 data_clean$Largest.Property.Use.Type == "Office" |
                                 data_clean$Largest.Property.Use.Type == "Hotel" |
                                 data_clean$Largest.Property.Use.Type == "Non-Refrigerated Warehouse"),]
data_clean$Largest.Property.Use.Type <- factor(data_clean$Largest.Property.Use.Type)

str(data_clean)
```

**Removing out lier from continuous variables and replacing missing values with average values**
```{r}
#Plot box plot for each continuous variable
par(mfrow=c(4,4))
ncol(data_clean)
for (i in 4:19){
  boxplot(data_clean[,i], xlab = colnames(data_clean)[i])
}
par(mfrow=c(1,1))
```
```{r}
#Find absolute value of z-score for each value in each column
df <- data_clean[,4:19]
z_scores <- as.data.frame(sapply(df, function(df) (abs(df-mean(df, na.rm = TRUE))/sd(df, na.rm = TRUE))))
data_clean <- data_clean[!rowSums(z_scores>3, na.rm = TRUE), ]

for (i in 1:19){
  if(length(which(is.na(data_clean[,i]))=="TRUE")>0){
  data_clean[which(is.na(data_clean[,i])),i] <- median(data_clean[,i], na.rm = TRUE)
  }
}
rm(i)
rm(df)
```

**Correlation Matrix for feature selection**
```{r}
cor_matrix <- data_clean
data_clean.cor <- cor(cor_matrix[4:19])
data_clean.cor.plot <- cor(cor_matrix[4:19])
colnames(data_clean.cor.plot) <- substring(colnames(data_clean[,4:19]),1,15)
rownames(data_clean.cor.plot) <- substring(colnames(data_clean[,4:19]),1,15)
corrplot(data_clean.cor.plot)

highlyCorrelated <- findCorrelation(data_clean.cor, cutoff=0.75, names = TRUE)

#Removing highly correlated columns
data_clean <- data_clean[,-which(colnames(data_clean) %in% highlyCorrelated)]
```

**Renaming columns**
```{r}
length(colnames(data_clean))
names_col <- c("PostalCode","LargeUse", "SubmissionStatus",
                          "BuiltYear","CountofBuildings", "Occupancy", 
                          "WeatherNormSiteEUI", "WeatherNormElecEUI", "WeatherNormNGEUI",
                          "ElecUse","DirectGHG","WaterIntensity","EnergyScore")
length(names_col)
colnames(data_clean) <- names_col
rm(names_col)
```

```{r}
#Remove duplicate entries
data_clean <- unique(data_clean)

str(data_clean)
```

```{r}
#Visualizing the distribution of Target variable for any skewed distribution
barplot(data_clean$EnergyScore)
```

**Split data in train and test in 70-30**
```{r}
dt <- sort(sample(nrow(data_clean), nrow(data_clean)*.7))

train_data <- data_clean[dt,]
test_data <- data_clean[-dt,]
rm(dt)
```

**scaling variables for uniform range. As the range for EnergyScore is not much, we will scale only the X variables and not target variable**
```{r}
train_data[,4:12] <- scale(train_data[,4:12])
test_data[,4:12] <- scale(test_data[,4:12])
```

```{r}
#Removing the Submission Status column in train data. This column can later be used to predict only 'Not Submitted' Entries in test data

train_data_submission_col <- train_data$SubmissionStatus
train_data$SubmissionStatus <- NULL

test_data_submission_col <- test_data$SubmissionStatus
test_data$SubmissionStatus <- NULL

str(train_data)
str(test_data)
```

**Modeling with all variable and understanding each variable's significance**
```{r}
#10 fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "RMSE"
```

```{r}
#Simple rpart - decision tree
set.seed(5)
fit.rpart <- train(EnergyScore~., data=train_data, method="rpart",metric = "RMSE", 
                   trControl=control, tuneLength = 15)
fit.rpart
#MAE with decision tree = 13
```

```{r}
#xgboost
#set.seed(7)
#fit.boost <- train(EnergyScore~., data=train_data, method="xgbTree")
#fit.boost

#knn
#This algorithm is not run due to resource limitations
#set.seed(5)
#fit.knn <- train(EnergyScore~., data=train_data, method="knn")

#random forest
#This algorithm is not run due to resource limitations
#set.seed(5)
##fit.rf <- train(EnergyScore~., data=train_data, method="rf")
```

```{r}
#Based on the comparative results of 2 algorithms (MAE and r-squared), we can go ahead with xGBoost 
#Further the accuracy of the selected model can be improved with Cross validation and tuning the hyper parameters for XGBoost
#It is not completed in this exercise due to resource limitations on the personal computer
#The model can be further improved by removing the non-important variables as follows. This assignment does not cover it due to time constraints.
Importance <- varImp(fit.rpart, scale=FALSE)
plot(Importance, top = 10)
```

**Test the data with selected model**
```{r}
predictions <- predict(fit.rpart, test_data[,1:11])
MAE(test_data$EnergyScore,predictions)
```

**Observations**
```{r}
#MAE for train data = 13
#MAE for test data = 19
#As test model MAE is about 2 times the MAE of train model, we can understand the model is overfit 
#We can perform k-fold cv to tune the parameters for xGBoost and run random forest for further improvement in the trained model
#We will skip this step due to time limitation and resource limitations
  ```

**Improving the model accuracy (difference between the MAE of train and test and MAE of test itself)**
```{r}
#Build dataset with important variable contributing more than 95% (drop postal code variable)
train_data_mod <- train_data[,2:12]
test_data_mod <- test_data[,2:12]

train_data_mod$SubmissionStatus <- NULL
test_data_mod$SubmissionStatus <- NULL

set.seed(7)
fit.rpart.mod <- train(EnergyScore~., data=train_data_mod, method="rpart")
fit.rpart.mod

#set.seed(7)
#fit.boost.mod <- train(EnergyScore~., data=train_data_mod, method="xgbTree")
#fit.boost.mod

predictions.mod <- predict(fit.rpart.mod, test_data_mod[,1:10])
MAE(test_data_mod$EnergyScore,predictions)
```

**Conclusion - We can derive a better regression model by running xGboost and tuning the parameters further to reduce overfitting.**

**We can explore clustering approach for possible accuracy improvement by creating similar subsets by location (weather) and building demography**

**The End**

